Hybrid platforms and bargaining power

### 2024-05-09 src/paper/application.tex
Update extension 1

\subsection{Platform has lower innate bargaining power}
\label{sec:lower_bargaining_power}

In the previous sections, I assumed that the platform and the fringe firms split total profits according to their Shapley values.
Because of the symmetry property of the Shapley value, this excludes the possibility of players having some kind of innate bargaining power, which is not related to their profit functions.
In this extension, I will relax this assumption, and assume instead that profits are shared according to \emph{weighed} values.
These are a generalization of the Shapley value, where each player has a weight $w_i$ that can, in certain settings, be thought of as a parameter describing bargaining power.\footnote{
    \textcite{hart1996bargaining} and \textcite{stole1996intra} provide foundations for this interpretation.
}
TODO: rewrite this
% Similarly to the standard Shapley value, one can show that outcomes based on weighted values also satisfy \cref{ass:profit_sharing}.
% Specifically, when the weight of the platform is $\lambda_P$, the functions $w_P$ and $w_F$ are given by
% \begin{align*}
%     w_P(s) &= \lambda_P s ^ {\lambda_P - 1} \\
%     w_F(s) &= s ^ {\lambda_P}.
% \end{align*}

\textcolor{red}{TODO: can I show that \cref{ass:single_crossing} still holds for any $\lambda_P$? If not, focus on a specific case, such as the one on the graphs.}

The example parametrizatrion is the same as in the main model, with the only difference that the platform has a lower innate bargaining power ($\lambda_P = 0.5$).
As the conditions in \cref{ass:single_crossing} are also satisfied in this case, the main observations from \cref{sec:results} also apply.
That is, as shown on \cref{fig:fringe_entry_eq_low_lambda}, total aggregate is decreasing in $N_P$ throughout the hybrid regime, due to the platform's products displacing more fringe products than their total number.

The main difference is that, due to the lower bargaining power of the platform, the implied entry fee (\cref{fig:platform_profits_low_lambda}) is lower than the benchmark, unilaterally set one for low enough $N_P$.
As a consequence, total aggregate, and thus consumer surplus, is also above the benchmark outcome in that case.
Furthermore, there exists some $\bar{N}_P$ for which the unilateral price setting and bargaining outcomes coincide in terms of all variables, due to the implied entry fee being the same.

Another major difference pertains to the total profits of the platform as a function of $N_P$.
It is still true that they are increasing in the number of the platform's products, but now this increase is even more pronounced for low values of $N_P$.
The reason is that for low $N_P$, the implied entry fee is lower than the optimal one.
Furthermore, it increases in $N_P$ as a result of the platform's increased bargaining power.
Therefore, the platform's profits are increasing in $N_P$ even more than in the benchmark case or the main model, up to the point where the implied entry fee is equal to the optimal one.
After that, the same effect as in the main model kicks in, and the increase in profits is lower than in the benchmark case.

\Cref{fig:equilibrium_low_lambda_entry_fees} demonstrates an important consequence of this phenomenon.
Imagine that the platform can decide on the number of its products at time 0, at an investment cost of $I_P$ per product.
In the main model, due to the implied entry fee being higher than optimal even for $N_P = 0$, hybrid operation is never optimal.
The platform would always prefer either pure retail or pure marketplace operation.
In contrast, when the platform has lower bargaining power, there exists a range of $I_P$ values (e.g. $I_P = 0.2$) for which the platform would prefer to operate in hybrid mode.
Furthermore, if hybrid operation is not allowed, the platform might switch to become a pure marketplace, thereby increasing consumer surplus.

@@

                \addplot[draw=none,name path=hybrid] table[x=N_P,y=hybrid]{\equilibrium};
                \addplot[draw=none,name path=hybrid_alt] table[x=N_P,y=hybrid]{\equilibriumProfitOneSidedLowLambda};
                \addplot[draw=none,name path=bottom] {0};
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid and bottom];
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid_alt and bottom];
            \end{axis}
        \end{tikzpicture}
        \caption{Equilibrium number of fringe firms ($N_F$)}
        \label{fig:fringe_entry_eq_low_lambda}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    xlabel={$N_P$},
                    xmin=0, xmax=1.5, % adjust these values as needed
                    ymin=0, ymax=3, % adjust these values as needed
                    width=\linewidth, % adjust the width of the plot
                    axis lines=left,
                    xtick={0, 1.5},
                    legend pos=south east,
                ]
                \addplot[color=myblue,mark=none,thick] table[x=N_P,y=A_bench]{\equilibrium};
                \addplot[color=myred,mark=none,thick] table[x=N_P,y=A]{\equilibrium};
                \addplot[color=mygreen,mark=none,thick] table[x=N_P,y=A]{\equilibriumProfitOneSidedLowLambda};
                \legend{Benchmark, Bargaining ($\lambda = 1$), Bargaining ($\lambda = 0.5$)}

                \addplot[draw=none,name path=hybrid] table[x=N_P,y=hybrid]{\equilibrium};
                \addplot[draw=none,name path=hybrid_alt] table[x=N_P,y=hybrid]{\equilibriumProfitOneSidedLowLambda};
                \addplot[draw=none,name path=bottom] {0};
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid and bottom];
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid_alt and bottom];
            \end{axis}
        \end{tikzpicture}
        \caption{Equilibrium number of fringe firms ($N_F$)}
    \end{subfigure}
    \caption{Equilibrium outcomes in the case when the platform has lower innate bargaining power ($N_C = 1, V_P = 1, V_F = 1, I_F = 0.2$). As in every other case, the platform's profits are increasing in $N_P$, while the fringe's profits are decreasing. Furthermore, consumer surplus is also decreasing as a result in a decrease in total product variety. The shaded region represents hybrid operation, while in the unshaded region, the platform operates in pure retail mode.}
    \label{fig:equilibrium_low_lambda}
\end{figure}

>>>>

\subsection{Playrs have different innate bargaining power}
\label{sec:lower_bargaining_power}

In the previous sections, I assumed that the platform and the fringe firms split total profits according to their Shapley values.
Because of the symmetry property of the Shapley value, this excludes the possibility of players having some kind of innate bargaining power, which is not related to their profit functions.
In this extension, I will relax this assumption, and assume instead that profits are shared according to \emph{weighed} values.
These are a generalization of the Shapley value, where each player has a weight $w_i$ that can, in certain settings, be thought of as a parameter describing bargaining power.\footnote{
    \textcite{hart1996bargaining} and \textcite{stole1996intra} provide foundations for this interpretation.
}
I still assume that fringe firms are identical, also in regard to their bargaining weights.
Therefore, the only difference is between the platform and the fringe firms.
Let us denote the platform's bargaining weight as $\lambda_P \in \mathbb{R}^+$, and without loss of generality, normalize the fringe firms' weights to 1.

As shown in \cref{sec:cooperative_game_weighted}, the weighted Shapley value, and thus final profits in this setting is given by
\begin{proposition}
    \label{prop:weighted_shapley_value}
    The Shapley value of the platform and the fringe firms, and thus final total profits, are given by:
    \begin{align*}
        \pi^t_P(N_P, N_F) &= N_C \int_0^1 \lambda_P s ^ {\lambda_P - 1} \Pi(N_P,s N_F) \ds, \\
        \pi^t_F(N_P, N_F) &= N_C \int_0^1 s ^ {\lambda_P} \partial_2 N_F \Pi(N_P,s N_F) \ds.
    \end{align*}
\end{proposition}
It is similar to the non-weighted value in that it is an average of marginal contributions, but now those averages are weighted, with the weight function depending on platform's bargaining weight.

The interpretation of $\lambda_P$ as a bargaining weight for the platform is supported by the fact that the platform's profit function is increasing in it.
\begin{proposition}
    \label{prop:lambda_P_comparative}
    For any fixed $N_F >0, N_P \geq 0$, the platform's total profits are increasing in its bargaining weight $\lambda_P$, while the fringe's profits are decreasing in it:
    \begin{align*}
        \frac{\partial \pi_P^t(N_P, N_F)}{\partial \lambda_P} &> 0, \\
        \frac{\partial \pi_F^t(N_P, N_F)}{\partial \lambda_P} &< 0.
    \end{align*}
\end{proposition}
In fact, the limits as $\lambda_P \to 0$ and $\lambda_P \to \infty$ are quite intuitive: in the former case, the platform's profits are zero, while in the latter case it is able to appropriate all profits.
That is, these limits correspond to the platform either receiving or making take-it or leave-it offers.

\textcolor{red}{TODO: can I show a result analogous to \cref{lem:shape_of_fringe_profit} still holds for any $\lambda_P$? If not, focus on a specific case, such as the one on the graphs.}

The example parametrizatrion is the same as in the main model, with the only difference that the platform has a lower innate bargaining power ($\lambda_P = 0.5$).
The main result, namely that the platform's product variety has a negative effect on consumer welfare in the hybrid regime, still holds.
That is, as shown on \cref{fig:fringe_entry_eq_low_lambda}, total aggregate is decreasing in $N_P$ throughout the hybrid regime, due to the platform's products displacing more fringe products than their total number.

The main difference is that, for low enough $N_P$, the implied entry fee (\cref{fig:platform_profits_low_lambda}) is lower than the benchmark, unilaterally set one,  due to the lower bargaining power of the platform.
As a consequence, total aggregate, and thus consumer surplus, is also above the benchmark outcome in that case.
Furthermore, there exists some $\bar{N}_P$ for which the unilateral price setting and bargaining outcomes coincide in terms of all variables, due to the implied entry fee being the same.

Another major difference pertains to the total profits of the platform as a function of $N_P$.
It is still true that they are increasing in the number of the platform's products, but now this increase is even more pronounced for low values of $N_P$.
The reason is that for low $N_P$, the implied entry fee is lower than the optimal one.
Furthermore, it is increasing in $N_P$ as a result of the platform's increased bargaining power.
Therefore, the platform's profits are increasing in $N_P$ even more than in the benchmark case or the main model, up to the point where the implied entry fee is equal to the optimal one.
After that, the same effect as in the main model kicks in, and the increase in profits is lower than in the benchmark case.

\Cref{fig:equilibrium_low_lambda_entry_fees} demonstrates an important consequence of this phenomenon.
Imagine that the platform can decide on the number of its products at time 0, at an investment cost of $I_P$ per product.
In the example for the main model, due to the implied entry fee being higher than optimal even for $N_P = 0$, hybrid operation is never optimal.
The platform would always prefer either pure retail or pure marketplace operation.
In contrast, when the platform has lower bargaining power, there exists a range of values for $I_P$ (most notably, $I_P = I_F = 0.2$) for which the platform would prefer to operate in hybrid mode.
Furthermore, if hybrid operation is not allowed, the platform might switch to become a pure marketplace, thereby increasing consumer surplus.

@@

                \addplot[draw=none,name path=hybrid] table[x=N_P,y=hybrid]{\equilibrium};
                \addplot[draw=none,name path=hybrid_alt] table[x=N_P,y=hybrid]{\equilibriumProfitOneSidedLowLambda};
                \addplot[draw=none,name path=bottom] {0};
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid and bottom];
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid_alt and bottom];
            \end{axis}
        \end{tikzpicture}
        \caption{Equilibrium number of fringe firms ($N_F$)}
        \label{fig:fringe_entry_eq_low_lambda}
    \end{subfigure}
    \hfill
    \begin{subfigure}[b]{0.45\textwidth}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    xlabel={$N_P$},
                    xmin=0, xmax=1.5, % adjust these values as needed
                    ymin=0, ymax=3, % adjust these values as needed
                    width=\linewidth, % adjust the width of the plot
                    axis lines=left,
                    xtick={0, 1.5},
                    legend pos=south east,
                ]
                \addplot[color=myblue,mark=none,thick] table[x=N_P,y=CS_bench]{\equilibrium};
                \addplot[color=myred,mark=none,thick] table[x=N_P,y=CS]{\equilibrium};
                \addplot[color=mygreen,mark=none,thick] table[x=N_P,y=CS]{\equilibriumProfitOneSidedLowLambda};
                \legend{Benchmark, Bargaining ($\lambda = 1$), Bargaining ($\lambda = 0.5$)}

                \addplot[draw=none,name path=hybrid] table[x=N_P,y=hybrid]{\equilibrium};
                \addplot[draw=none,name path=hybrid_alt] table[x=N_P,y=hybrid]{\equilibriumProfitOneSidedLowLambda};
                \addplot[draw=none,name path=bottom] {0};
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid and bottom];
                \addplot [fill=black, fill opacity=0.05] fill between [of=hybrid_alt and bottom];
            \end{axis}
        \end{tikzpicture}
        \caption{Equilibrium consumer surplus ($CS$)}
    \end{subfigure}
    \caption{Equilibrium outcomes in the case when the platform has lower innate bargaining power ($N_C = 1, V_P = 1, V_F = 1, I_F = 0.2$). As in every other case, the platform's profits are increasing in $N_P$, while the fringe's profits are decreasing. Furthermore, consumer surplus is also decreasing as a result in a decrease in total product variety. The shaded region represents hybrid operation, while in the unshaded region, the platform operates in pure retail mode.}
    \label{fig:equilibrium_low_lambda}
\end{figure}
